---
title: "EDS241: Assignment 4"
author: "Halina Do-Linh"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: no
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)
# load packages
librarian::shelf(here,
                 tidyr,
                 stringr,
                 janitor,
                 estimatr,
                 cowplot,
                 ggplot2,
                 tinytex,
                 datasets,
                 tibble,
                 tidyverse,
                 huxtable,
                 car,
                 readxl,
                 AER) # AER package has ivreg()

# devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation
```

\noindent For Assignment 4, we will estimate the price elasticity of
demand for fresh sardines across 56 ports located in 4 European
countries with monthly data from 2013 to 2019. The data are contained in
the file `EU_sardines.csv`.

\noindent We are using the following variables: `year`, `month`,
`country`, `port` (port where sardines are landed and sold),
`price_euro_kg` (price per kg in €), and `volume_sold_kg` (quantity of
sardines sold in kg). In the questions below, I use `log()` to denote
the natural logarithm.

# Load Data

```{r}
sardines_data <- read_csv(here::here("assignments/data/EU_sardines.csv"))
```

# Question A

```{r}
# log of volume sold and price 
sardines_log <- sardines_data %>% 
  mutate(price_euro_kg_log = log(price_euro_kg),
         volume_sold_kg_log = log(volume_sold_kg))

# bivariate regression
model_1 <- lm_robust(data = sardines_log, 
                     volume_sold_kg_log ~ price_euro_kg_log)
huxreg(model_1)

# test null hypothesis that price elasticity = -1
linearHypothesis(model_1, c("price_euro_kg_log = -1"), white.adjust = "hc2")
```

\noindent \textbf{Answer:} The price elasticity of demand for sardines `r round(model_1$coefficients[2], 2)`. And we reject the null hypothesis that price elasticity is equal to -1 because the p-value is statistically significant. 

\newpage

# Question B

```{r}
# first stage regression
model_2 <- lm_robust(data = sardines_log, price_euro_kg_log ~ wind_m_s)
huxreg(model_2)
```

\noindent \textbf{Answer:} The estimated coefficient $\beta\_1$ on `wind_m_s` tells us that for each 1 $m/s$ of wind speed the price of sardines in euros increases by `r round(model_2$coefficients[2], 3)`. This is the expected sign because fishing conditions are worse when it is windy, meaning supply will likely go down causing prices to increase.

\newpage

```{r}
# instrument relevance and non--weak test 
# want f-statistic to be greater than 10
hypo_1 <- linearHypothesis(model_2, # first stage regression model
                 c("wind_m_s = 0"), # null hypothesis
                 white.adjust = "hc2") # robust heteroskedastic errors
```

\noindent \textbf{Answer:} The F-statistic is `r round(hypo_1$Chisq[2], 2)` and is greater than 10 and is therefore not a weak instrument.

# Question C

```{r}
tsls1 <- ivreg(volume_sold_kg_log ~ price_euro_kg_log | wind_m_s, data = sardines_log)
huxreg(tsls1)
```

\noindent \textbf{Answer:} The estimated price elasticity of demand for sardines is `r  round(tsls1$coefficients[2], 2)`. 

# Question D

```{r}
# tsls with fixed effects
tsls_fixed <- ivreg(volume_sold_kg_log ~ price_euro_kg_log + 
                      as.factor(country) + as.factor(year) + as.factor(month) | 
                      wind_m_s + 
                      as.factor(country) + as.factor(year) + as.factor(month), 
                    data = sardines_log)
```


```{r}
# need to use the regression from the first stage with the fixed effects
model_fixed <- lm_robust(data = sardines_log, price_euro_kg_log ~ wind_m_s +
                           as.factor(country) + as.factor(year) + as.factor(month))

hypo_fixed <- linearHypothesis(model_fixed, c("wind_m_s=0"), white.adjust = "hc2")
```

\noindent \textbf{Answer:} The estimated price elasticity of demand is `r round(tsls_fixed$coefficients[2], 2)`. And the F-statistic testing for wind as a relevant and non-weak instrument is `r round(hypo_fixed$Chisq[2], 2)`.
