---
title: "Notes"
author: "Halina Do-Linh"
date: "3/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Notes


## Question A
- Best way to test null hypo? Is CI's enough? 
- double check model

```{r}
# huxreg(model_1, 
#        bold_signif = 0.05,
#        ci_level = 0.95,
#        error_format = ("({conf.low} -- {conf.high})"))
```


We can reject the null hypothesis that price elasticity is equal to -1 because the confidence intervals for `price_euro_kg_log` are within the range `r round(model_1$conf.low[2], 3)` and `r round(model_1$conf.high[2], 3)` at a 95% significance level.

## Questions
- How to know what is the expected sign? You catch less when it's bad conditions aka windy so we expect the sign to be positive therefore more expensive price for sardines when wind increases

## Question
- What is the correct model? Where to put the fixed effects?
- What is over identified model mean? I.e. from IV.R lab
- Do we need a different model for getting the F-statistic? 


```{r}
# tsls with fixed effects
tsls_fixed <- ivreg(volume_sold_kg_log ~ price_euro_kg_log + as.factor(country) + as.factor(year) + as.factor(month) | as.factor(country) + as.factor(year) + as.factor(month) + wind_m_s, data = sardines_log)

# need to have the same covariates in the same 
# second stage | first stage
# left out is the instrument and the endogenous 
```

```{r}
# instrument relevance and non--weak test 
# want f-statistic to be greater than 10
# need to use the regression from the first stage with the fixed effects
model_fixed <- lm_robust(data = sardines_log, price_euro_kg_log ~ wind_m_s + as.factor(country) + as.factor(year) + as.factor(month))

hypo_fixed <- linearHypothesis(model_fixed, 
                 c("wind_m_s=0"), 
                 white.adjust = "hc2")

## get an error here....which model should I use?
# test for strong and weak instrument only happens in the first stage regression, not in the second stage
```

\begin{align}
  $price_euro_kg_log_{i} = \beta_0 + \beta_1 wind_m_s_{1i} + u_{i}$
\end{align}

```{r}
# instrument relevance and non--weak test 
# want f-statistic to be greater than 10
# need to use the regression from the first stage with the fixed effects
model_fixed <- lm_robust(data = sardines_log, price_euro_kg_log ~ wind_m_s + as.factor(country) + as.factor(year) + as.factor(month))

hypo_fixed <- linearHypothesis(model_fixed, c("wind_m_s=0"), white.adjust = "hc2")
```
