---
title: "Notes HW 1"
author: "Halina Do-Linh"
date: "1/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Notes

For Question E from TA OH

How does robust SD error vary from regular? robust does HC2 a more
conservative measurement of standard error HC2 regression estimation
part and the inference part is this statistically significant or not
coefficient: 

- is this beta linked positively or negative? 
- does this make y or our prediction go up or down 
- make inferential terms based on standard errors

p value, confident intervals, and standard errors are all related

standard error gives us context about p value. so be more suspicious of
our p value if our standard errors be large?

picking the sig level is what makes p value are what arbitrary

standard errors as significant: reports automatically little sig stars
which directly interprets our p values

hetero robust stand errors auto adjusts our standard errors

low birth are increased as pm 2.5 increases

# Question from hw 1 - Question E

Suppose a new air quality policy is expected to reduce PM2.5 concentration by 2
micrograms per cubic meters. Predict the new average value of LowBirthWeight and
derive its 95% confidence interval. Interpret the 95% confidence interval

# Question E

\noindent The code chunk below shows how to produce a regression with a 95% confidence interval when `PM25` is now changing by -2.

need to manipulate dataframe to  create a pm2_5 policy 

notes from TA OH:
- don't want to reduce everything by 2 because then we get the same coefficent (which we did)
- use what we already estimated and use those coefficients to make a statement about a potential policy
- what is the predicted mean effect of reducing PM2.5 by 2 micrograms/cubic meter? ANS: we know what this looks like because it's a linear model. we know that by doubling that it would reduce more or the slope would be steeper? 

minus 2 times beta coefficient 

intercept is not in the confidence interval equation 

What equation are we using? ANS: slide 13 on lecture 3/4
beta truee effect = estimated y - estimated 

```{r}
# # create new df with new policy
# ces_policy <- data_ces_clean %>% 
#   mutate(pm2_5_policy = pm2_5 - 2)
# 
# # predict mean of new low birth weight
# # y = mx + b
# ces_policy <- ces_policy %>% 
#   mutate(low_birth_weight_policy = (model1$coefficients[2] * pm2_5_policy) + model1$coefficients[1])
#   
# lbw_policy = mean(ces_policy$low_birth_weight_policy)
# 
# model2 <- lm_robust(formula = low_birth_weight_policy ~ pm2_5_policy, data = ces_policy)
# huxreg(model2)
# 
# confint(object = model2, parm = "pm2_5_policy", level = 0.95)
```

\noindent **Answer: PM 2.5 did not change....TBD**

Note: How to derive the confidence interval? 